Reproducible Research Peer Assessment 1
========================================================
## What is the mean total number of steps taken per day?

#### 1. Calculate the total number of steps taken per day.

```{r}
activity <- read.csv ("~/Downloads/activity.csv")
library(plyr)
activity1<-na.omit(activity)
stepsday<- ddply(activity1,.(date),summarize,steps=sum(steps))
stepsday
```

#### 2. Make a histogram of the total number of steps taken each day.

```{r}
histogram <- hist(stepsday$steps,breaks=50,main = "Histogram of Total Steps Per Day (Breaks = 50)", xlab = "Steps per day")
```

#### 3. Calculate and report the mean and median of the total number of steps taken per day.

```{r}
meanvalue <-mean(stepsday$steps,na.rm=TRUE)
medianvalue <- median(stepsday$steps, na.rm=TRUE)
```

The mean of the steps taken per day is `r meanvalue` and the median is `r medianvalue`.

## What is the average daily activity pattern?

#### 1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, average across all days(y-axis).

```{r}
activity1 <- na.omit(activity)
int.steps <- ddply(activity1,.(interval),summarize,steps=mean(steps))
plotinterval <- plot(int.steps$interval,int.steps$steps,type="l",main ="Average steps per 5-min interval across all days", xlab = "5 min interval across the day", ylab = "Average Steps Taken in Interval")
plotinterval
```

#### 2. Which 5-min interval, on average across all days in the dataset, contains the maximum number of steps.

```{r}
int.steps <- arrange(int.steps,desc(steps))
max_int <-int.steps$interval[1]
max_int
```

The maximum number of steps on average occurred in the `r max_int` interval.

## Inputing missing values

#### 1. Calculate and report the total number of missing values in the dataset.
```{r}
NAValues <- sum(is.na(activity))
NAValues
```

The total numnber of missing values in the dataset is `r NAValues`.

#### 2. Devise a strategy for filling in all the missing values in the dataset.

All NAs will be replaced with the interval's mean value.

#### 3. Create a new dataset that is equal to the original set but with mssing data filled in.
```{r}

## Create additional column to show average of the interval.

newactivity <-merge(activity,int.steps, by="interval")

for (i in 1:nrow(activity)) {
  if (is.na(newactivity$steps.x[i])) {
    newactivity$steps.x[i] <- newactivity$steps.y[i]
  }
}

head(newactivity)
```

#### 4. Create a new historgram of the total number of steps taken each day. Calculate and report the mean and median total number of steps taken per day. Do the values differ from the first part of the assignmet? What is the impact of imputtng missing data on the estimates of the total daily number of steps?

```{r}
newstepsday<- ddply(newactivity,.(date),summarize,steps=sum(steps.x))
new_histogram <- hist(newstepsday$steps,breaks=50,main = "(NEW) Histogram of Total Steps Per Day (Breaks = 50)", xlab = "Steps per day")
new_histogram
newmeanvalue <-mean(newstepsday$steps)
newmedianvalue <- median(newstepsday$steps)
diffmeanvalue <- newmeanvalue-meanvalue
diffmedianvalue <- newmedianvalue-meanvalue
```

The new mean steps taken per day is `r newmeanvalue` which is `r diffmeanvalue` from the original value.

The new median steps taken per day is `r newmedianvalue` which is `r diffmedianvalue` from the original value.

## Are there differences in activity patterns between weekdays and weekends?

#### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
newactivity$date <- as.Date(newactivity$date, "%Y-%m-%d")
newactivity$day<- weekdays(newactivity$date)
for (i in 1:nrow(newactivity)) {
  if (newactivity$day[i]=="Monday") {
    newactivity$day[i] <- "Weekday"
  }
  if (newactivity$day[i]=="Tuesday") {
    newactivity$day[i] <- "Weekday"
  }
   if (newactivity$day[i]=="Wednesday") {
    newactivity$day[i] <- "Weekday"
  }
   if (newactivity$day[i]=="Thursday") {
    newactivity$day[i] <- "Weekday"
  }
   if (newactivity$day[i]=="Friday") {
    newactivity$day[i] <- "Weekday"
  }
   if (newactivity$day[i]=="Saturday") {
    newactivity$day[i] <- "Weekend"
  }
   if (newactivity$day[i]=="Sunday") {
    newactivity$day[i] <- "Weekend"
  }
}
head(newactivity)
```

#### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
weekend<-subset(newactivity, newactivity$day =='Weekend')
weekday<-subset(newactivity,newactivity$day =='Weekday')

weekday_int <- ddply(weekday,.(interval),summarize,steps=mean(steps.x))
weekday_int$day <- "weekday"
weekend_int <- ddply(weekend,.(interval),summarize,steps=mean(steps.x))
weekend_int$day <- "weekend"
new_activity_day <- rbind(weekday_int, weekend_int)
library(lattice)

plot <- xyplot(new_activity_day$steps ~ new_activity_day$interval | new_activity_day$day, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")

plot

